### 2021-07-23 paper/tex/paper.tex
ensure all stats are up to date

\documentclass[twocolumn, twocolappendix]{aastex63}
\received{\today}
\shorttitle{Detecting DCOs with LISA}
\graphicspath{{../figures/}}

\usepackage{lipsum}
\usepackage{physics}
\usepackage{multirow}
\usepackage{xspace}
\usepackage{natbib}
\usepackage{fontawesome5}
\usepackage{xcolor} % to add more color options
\usepackage{wrapfig}

% remove indents in footnotes
\usepackage[hang,flushmargin]{footmisc}

\newcommand{\todo}[1]{{\color{red}{[TODO: #1}]}}
\newcommand{\tom}[1]{\textcolor{ForestGreen}{[Tom: #1]}}
\newcommand{\floor}[1]{\textcolor{Bittersweet}{[Floor: #1]}}
% custom function for adding units
\makeatletter
\newcommand{\unit}[1]{%
    \,\mathrm{#1}\checknextarg}
\newcommand{\checknextarg}{\@ifnextchar\bgroup{\gobblenextarg}{}}
\newcommand{\gobblenextarg}[1]{\,\mathrm{#1}\@ifnextchar\bgroup{\gobblenextarg}{}}
\makeatother

\newcommand{\avg}[1]{\left\langle#1\right\rangle}

%% Physics variations shortcuts
\newcommand{\modFid}{A}
\newcommand{\modBetaLow}{B}
\newcommand{\modBetaMed}{C}
\newcommand{\modBetaHigh}{D}
\newcommand{\modCaseBB}{E}
\newcommand{\modCaseBBOpt}{F}
\newcommand{\modAlphaLowest}{G}
\newcommand{\modAlphaLow}{H}
\newcommand{\modAlphaHigh}{I}
\newcommand{\modAlphaHighest}{J}
\newcommand{\modOpt}{K}
\newcommand{\modRapid}{L}
\newcommand{\modNSLow}{M}
\newcommand{\modNSHigh}{N}
\newcommand{\modNoPISN}{O}
\newcommand{\modSigLow}{P}
\newcommand{\modSigLower}{Q}
\newcommand{\modNoBH}{R}
\newcommand{\modWRLow}{S}
\newcommand{\modWRHigh}{T}

\newcommand{\modRangeMT}{B-F}
\newcommand{\modRangeCE}{G-K}
\newcommand{\modRangeSN}{L-R}
\newcommand{\modRangeML}{S-T}

\newcommand{\nModels}{20}
\newcommand{\nMinusOneModels}{19}

\newcommand{\BHBHFourYear}{74}
\newcommand{\BHNSFourYear}{42}
\newcommand{\NSNSFourYear}{8}
\newcommand{\BHBHTenYear}{TODO}
\newcommand{\BHNSTenYear}{TODO}
\newcommand{\NSNSTenYear}{TODO}

% number of systems that are definitely not WDWDs
\newcommand{\BHBHNotWDWDFour}{33}
\newcommand{\BHNSNotWDWDFour}{16}
\newcommand{\NSNSNotWDWDFour}{5}
\newcommand{\BHBHNotWDWDTen}{TODO}
\newcommand{\BHNSNotWDWDTen}{TODO}
\newcommand{\NSNSNotWDWDTen}{TODO}

% percentage of systems above Chandrasekhar mass
\newcommand{\BHBHAboveMaxWDWDFourPerc}{17}
\newcommand{\BHNSAboveMaxWDWDFourPerc}{21}
\newcommand{\NSNSAboveMaxWDWDFourPerc}{3}
\newcommand{\BHBHAboveMaxWDWDTenPerc}{TODO}
\newcommand{\BHNSAboveMaxWDWDTenPerc}{TODO}
\newcommand{\NSNSAboveMaxWDWDTenPerc}{TODO}

% number of systems that can be determined as either a BH or NS
\newcommand{\BHBHEitherBHOrNSFour}{11}
\newcommand{\BHNSEitherBHOrNSFour}{6}
\newcommand{\NSNSEitherBHOrNSFour}{3}
\newcommand{\BHBHEitherBHOrNSTen}{TODO}
\newcommand{\BHNSEitherBHOrNSTen}{TODO}
\newcommand{\NSNSEitherBHOrNSTen}{TODO}
\newcommand{\BHBHEitherBHOrNSFourPerc}{14}
\newcommand{\BHNSEitherBHOrNSFourPerc}{13}
\newcommand{\NSNSEitherBHOrNSFourPerc}{35}
\newcommand{\BHBHEitherBHOrNSTenPerc}{TODO}
\newcommand{\BHNSEitherBHOrNSTenPerc}{TODO}
\newcommand{\NSNSEitherBHOrNSTenPerc}{TODO}

% percentage of systems in lower mass gap
\newcommand{\BHBHLowerMassGap}{35\%}
\newcommand{\BHNSLowerMassGap}{38\%}

% percentage of systems with q > 0.8
\newcommand{\BHBHqAbovePointEight}{8\%}
\newcommand{\BHNSqAbovePointEight}{0\%}
\newcommand{\NSNSqAbovePointEight}{90\%}

% percentage of systems with e > 0.01
\newcommand{\BHBHNotCirc}{86\%}
\newcommand{\BHNSNotCirc}{47\%}
\newcommand{\NSNSNotCirc}{86\%}

% percentage of systems with e > 0.3
\newcommand{\BHBHHighlyEccentric}{22\%}
\newcommand{\BHNSHighlyEccentric}{8\%}
\newcommand{\NSNSHighlyEccentric}{12\%}

% percentage of systems with something other than 1 detectable harmonic
\newcommand{\BHBHMultipleHarmonicsFourPerc}{55}
\newcommand{\BHNSMultipleHarmonicsFourPerc}{28}
\newcommand{\NSNSMultipleHarmonicsFourPerc}{66}
\newcommand{\BHBHMultipleHarmonicsTenPerc}{TODO}
\newcommand{\BHNSMultipleHarmonicsTenPerc}{TODO}
\newcommand{\NSNSMultipleHarmonicsTenPerc}{TODO}

% percentage of systems with >1 harmonic and in disc
\newcommand{\BHBHEccInDiscFourPerc}{40}
\newcommand{\BHNSEccInDiscFourPerc}{25}
\newcommand{\NSNSEccInDiscFourPerc}{60}
\newcommand{\BHBHEccInDiscTenPerc}{TODO}
\newcommand{\BHNSEccInDiscTenPerc}{TODO}
\newcommand{\NSNSEccInDiscTenPerc}{TODO}

% percentage of systems with fractional chirp mass uncertainty < 1
\newcommand{\BHBHChirpMassMeasureable}{20}
\newcommand{\BHNSChirpMassMeasureable}{17}
\newcommand{\NSNSChirpMassMeasureable}{4}

\newcommand{\confinv}[3]{$#1${\raisebox{0.5ex}{\tiny$_{-#2}^{+#3}$}}}

\definecolor{Blush}{rgb}{0.87, 0.36, 0.51}
\newcommand{\SdM}[1]{{\color{Blush}{#1}}}

% Milky Way notation
\newcommand{\achem}{\ensuremath{[\alpha/\mathrm{Fe}]}}

% \color{white}
% \pagecolor{black}

\begin{document}

% selma's newest catchy one
\title{Gravitational wave sources in our Galactic backyard: Predictions for BH and NS binaries in LISA}

% selma suggested removing BHNS
% \title{Predictions for neutron star and black hole binaries with LISA}

% original
%\title{Predictions for detecting BHNS and other double compact objects binaries with LISA}

% affiliations
\newcommand{\cfa}{Center for Astrophysics | Harvard \& Smithsonian, 60 Garden Street, Cambridge, MA 02138, USA}
\newcommand{\mpa}{Max-Planck-Institut für Astrophysik, Karl-Schwarzschild-Straße 1, 85741 Garching, Germany}
\newcommand{\cca}{Center for Computational Astrophysics, Flatiron Institute, 162 Fifth Ave, New York, NY, 10010, USA}
\newcommand{\UvA}{Anton Pannekoek Institute for Astronomy \& Grappa, University of Amsterdam, Postbus 94249, 1090 GE Amsterdam, The Netherlands}

\author[0000-0001-6147-5761]{T. Wagg}
\affiliation{\cfa}
\affiliation{\mpa}

\author[0000-0002-4421-4962]{F. Broekgaarden}
\affiliation{\cfa}

\author[0000-0001-9336-2825]{S. E. de Mink}
\affiliation{\mpa}
\affiliation{\UvA}
\affiliation{\cfa}

\author[0000-0002-6411-8695]{N. Frankel}
\affiliation{Max Planck Institute for Astronomy, Königstuhl 17, D-69117 Heidelberg, Germany}

\author[0000-0001-5484-4987]{L.A.C. van Son}
\affiliation{\cfa}
\affiliation{\UvA}
\affiliation{\mpa}

\author[0000-0001-7969-1569]{S. Justham}
\affiliation{School of Astronomy \& Space Science, University of the Chinese Academy of Sciences, Beijing 100012, China}
\affiliation{\UvA}
\affiliation{\mpa}

\correspondingauthor{Tom Wagg}
\email{tomjwagg@gmail.com}

\begin{abstract}
    \input{sections/abstract.tex}
\end{abstract}

\keywords{LISA, gravitational waves, black hole, neutron star, binary}

\section{Introduction} \label{sec:intro}
\input{sections/introduction.tex}

\section{Method} \label{sec:method}
\input{sections/methods.tex}

\section{Results I - Fiducial} \label{sec:results}
\input{sections/results.tex}

\section{Results II - Model variations} \label{sec:variations}
\input{sections/variations.tex}

\section{Discussion} \label{sec:discussion}
\input{sections/discussion.tex}

\section{Comparison with previous studies}\label{sec:compare_studies}
\input{sections/comparisons.tex}

\section{Conclusion \& Summary} \label{sec:conclusion}
\input{sections/conclusion.tex}

\software{COMPAS (version 02.12.00) \url{http://github.com/TeamCOMPAS/COMPAS}. \citep{Stevenson+2017, Vigna-Gomez+2018, Broekgaarden+2019}, \texttt{Python} available from \url{python.org}, matplotlib \citep{Hunter+2007},  \texttt{NumPy} \citep{vanderWalt+2011}, 
Astropy \citep[\url{http://www.astropy.org}][]{AstropyCollaboration+2013, AstropyCollaboration+2018}.}

\acknowledgments{We thank Mike Lau, Floris Kummer, Eva Laplace, Rob Farmer, Katie Sharpe, Mathieu Renzo, Katie Brevik, Alberto Sesana, Valeriya Korol, Will Farr, Tyson Littenberg and the members of the COMPAS collaboration for insightful discussions. TW also thanks Terence Lovatt for his advice and help on improving an earlier draft of this project. This project was funded in part by the European Union’s Horizon 2020 research and innovation program from the European Research Council (ERC, Grant agreement No. 715063), and by the Netherlands Organization for Scientific Research (NWO) as part of the Vidi research program BinWaves with project number 639.042.728. We further acknowledge the Black Hole Initiative funded by a generous contribution of the John Templeton Foundation and the Gordon and Betty Moore Foundation.}

\clearpage

\bibliographystyle{aasjournal}
\bibliography{references}

\allowdisplaybreaks
\appendix
\input{sections/appendix.tex}

\end{document}