### 2021-10-30 paper/tex/sections/variations.tex

In this section we explore the effect of varying the uncertain assumptions governing the evolution of binary system and the formation of compact objects. For this we use the model variations presented in \citet{Broekgaarden+2021}.

We first discuss the robustness of our predictions for the number of detectable systems (Section~\ref{sec:detection_rate_analysis}). We then discuss examples of how the observable properties of the detectable systems, such as the distribution of masses and eccentricities, are affected and can potentially be used to probe about the physics of double compact object formation (Section~\ref{sec:property_variations}).

%We note that impact of physics variations cannot always be easily understood in simple terms as the evolutionary path of a progenitor of a detectable source involves multiple interactions stages. Providing full explanations is beyond our current scope, but we hope nevertheless that this section gives the reader an impression of the impact of the physics variations on the rates.

\subsection{Detection rates}\label{sec:detection_rate_analysis}
In Figure~\ref{fig:detection_rates}, we show the number of LISA detections as predicted by our simulations considering variations in the physical assumptions.  We show predictions for  BHBH, BHNS and NSNS systems in the top, middle and bottom panel respectively. All the rates and their uncertainties plotted in this figure are also provided in Table~\ref{tab:detection_rates}. 
In the sections that follow, we briefly explain the variations considered and discuss the most prominent trends.

\subsubsection{Efficiency of mass transfer}

The efficiency of mass transfer, i.e.\ the fraction of mass lost by the donor through Roche lobe overflow that is accreted by the companion, is poorly constrained and is considered as one of the main uncertainties in binary evolution \citep[e.g.\ ][]{deMink+2007}. In our fiducial model A, we use a prescription where the 
accretion rate onto stellar companions is regulated by their thermal timescale, i.e. the timescale on which a star can react to changes and restore thermal equilibrium \citep[see e.g.\ ][]{Schneider+2015}.

In models \modBetaLow{}-\modBetaHigh{}, we adopt a fixed value for the mass transfer efficiency, $\beta$, from $\beta=0.25$ up to 0.75, in cases of stable mass transfer onto a stellar companion. For accretion onto NS and BH we still assume that their accretion is limited to the Eddington rate.

Nearly all systems that can be detected form through channels where the very first interaction is stable mass transfer. Generally, higher mass transfer efficiencies lead to higher masses for the accreting stars, but also leads to wider orbits \citep{Soberman+1997, vanSon+2020}. 
Changing $\beta$ thus already affects the masses and orbital separation after the first interaction phase, which in turn changes the starting conditions and outcome of all subsequent interactions phases. This makes it complicated to fully understand the impact of varying $\beta$ in simple terms, but we can distinguish two main patterns for higher and lower mass systems respectively.

For the most massive progenitors, increasing $\beta$ leads to secondary stars that are so massive and luminous that they experience strong wind mass loss. This leads to further widening of the orbit and it limits the amount by which the secondaries expand. Both of these effects can prevent secondaries from filling their Roche lobe. This means that they cannot initiate the reverse interaction needed to shrink the binary system and eventually produce a detectable compact binary system. We indeed see that increasing the mass transfer efficiency leads to 
a decrease of the expected number of detections for BHBHs and BHNSs, which originate from the most massive progenitors.

For lower mass progenitors, which primarily produce NSNS systems, we find the opposite. Increasing $\beta$ leads to an increase of detectable systems. This is in part because the increased mass of the secondary stars allows stars that would have other wise ended their life as a WD, can now become massive enough to form a NS. The same effect also allows some NS progenitors to become massive enough to BH progenitors, but this effect is less strong. Furthermore, the changes in secondary mass and orbital widening, affects the number of systems for which the reverse interaction successfully ejects the envelope and shrinks the orbit. Indeed, the net effect is an increase of NSNS systems with increasing $\beta$.

\begin{figure*}[p]
    \centering
    \includegraphics[width=\textwidth]{fig9_dco_detections.pdf}
    \caption{The number of expected detections in the LISA mission for different DCO types and model variations. Error bars show the 1- (solid) and 2-$\sigma$ (dotted) Poisson uncertainties. An arrow indicates that the error bar extends to zero. The left axis and grid lines show the number of detections in a 4-year LISA mission and the right axis shows an approximation of the number of detections in a 10-year mission (we scale the axis by $\sqrt{T_{\rm obs}}$, see Table~\ref{tab:detection_rates} for exact rates). Each model is described in further detail in Table~\ref{tab:physics_variations} and details of the fiducial assumptions are in Section~\ref{app:fiducial_physics}. See also Fig.~\ref{fig:dco_relative_rates} and Sec.~\ref{sec:detection_rate_analysis} for a discussion. \href{https://github.com/TomWagg/detecting-DCOs-in-LISA/blob/main/paper/figures/fig9_dco_detections.pdf}{\faFileImage} \href{https://github.com/TomWagg/detecting-DCOs-in-LISA/blob/main/paper/figure_notebooks/detections.ipynb}{\faBook}.}
    \label{fig:detection_rates}
\end{figure*}

\subsubsection{Common-envelope evolution}\label{sec:detection_rate_CE_trends}

The common-envelope phase constitutes the most challenging phase in the evolution of interacting binary systems \citep{Ivanova+2013}. The debate concerns the conditions required for the onset of a common envelope phase and, if a common envelope phase occurs, what the outcome is. Rapid populations synthesis simulations such as ours approximate both questions in a crude way. We therefore consider several model variations.

\paragraph{The efficiency parameter $\alpha_{\rm CE}$} To estimate the outcome of a CE phase, we use a simple consideration of the binding energy and orbital energy  \citep{Webbink+1984}. Our fiducial model assumes a common-envelope efficiency parameter $\alpha_{\rm CE} = 1$ which can be interpreted as the case where all the energy liberated by shrinking the orbit it used in an optimal way to unbind the envelope \citep{Webbink+1984}. There have been many attempts to constrain this parameter using observations and more recently also using 3D simulations, but no consistent picture has emerged. Therefore we consider variations in this parameter.

In models \modAlphaLowest{}-\modAlphaHighest{} we alter the common-envelope efficiency parameter $\alpha_{\rm CE}$ to $0.1, 0.5, 2.0$ and $10.0$ respectively. Values smaller than 1 may represent cases where not all energy is used efficiently to unbind the envelope, for example when part of the energy escapes in the form of radiation or if part is used to impart additional kinetic energy in the ejecta. Values larger than 1 may represent cases where additional energy sources can be tapped into, such as for example jets powered by accretion. The variations can also be seen as a way to cover uncertainties in estimates for the binding energy itself.

Increasing $\alpha_{\rm CE}$ makes common-envelope ejection more efficient or, in other words, less orbital shrinkage is needed to successfully eject the envelope. This has two consequences. (1) Firstly, a larger fraction of systems avoids merging during the CE phase. This increases the overall number of double compact object binaries. (2) Secondly, the systems that survive the CE phase are wider, possibly too wide to become detectable as gravitational wave sources. So, while the first consequence favours the formation of double compact object systems, the second consequence disfavours the subset of those close enough to be detected.

These two opposing effect result in fine tuning situation. Only a very small subset of progenitor systems that have the right orbital parameters prior to the CE phase to successfully produce detectable systems. Changing  $\alpha_{\rm CE}$ moves and changes this window in the parameter space that successfully leads to the formation of detectable systems. How the number of the detectable systems changes depends on whether the relevant parameter space grows or shrinks and on how well the relevant part of the parameters space is populated. Fully unravelling these effects and how they interplay with the assumed star formation history is beyond our scope (and possibly not even of large relevance given the severe simplifications). We will limit the further discussions to simply stating the trends we observe.

We find that the BHBH rate peaks for $\alpha_{\rm CE} = 1$ (model \modFid{}) and reduces whether we increase \textit{or} decrease $\alpha_{\rm CE}$. The BHNS rate follows the same pattern as BHBHs although the value of $\alpha_{\rm CE}$ which maximises BHNSs seems to be between 1 and 2. In contrast, for NSNSs we find that increasing $\alpha_{\rm CE}$ (models \modAlphaHigh{} and \modAlphaHighest{}) results in significantly higher rates.

\paragraph{The ``optimistic'' CE treatment} We further explore a model variation introduced by \citet{Belczynski+2007} often referred to as the ``optimistic'' CE scenario. This model also considers donor stars that are on the Hertzsprung as the progenitors for CE events.

In agreement with other studies, we find that this treatment leads to a large increase formation rate of BHBHs and with it the number of systems that LISA can detect. This is because BHBHs originate from the most massive progenitors, which in expand very significantly during the Hertzsprung gap phase in our simulations. The progenitors are likely initiate interaction during this phase. Our predictions for the number of detectable systems increases by a factor 2. The less massive progenitors of BHNSs are less strongly affected, and for NSNS we do not see a significant increase.

\paragraph{Case BB mass transfer} For completeness and consistency with \citet{Broekgaarden+2021}, we have also included model variations that consider uncertainties in case BB mass transfer. his is a phase of mass transfer where the donor star has already lost its hydrogen envelope in a prior interaction, but fills its Roche lobe again as expands during helium shell burning phase \citep[e.g.][]{Dewi+2003, Vigna-Gomez+2018}. This is of particular interest for the formation of NS stars, as their lower mass progenitors are swelling the most during this phase \citet[e.g.][and references therein]{Laplace+2020}. Nearly all NSNS systems from through case BB mass transfer

In model \modCaseBB{} we effectively exclude any systems that evolve through case BB mass transfer. As expected, this kills nearly all NSNS systems. We also find a heavy reduction of the BHNS systems y a factor 6. The BHBH systems are not affected, as expected, since case BB mass transfer does not play a role for high mass progenitors.

Finally, in model {[\bf TOM.. please add label]} [=F] we consider a situation where case BB mass transfer always leads to unstable mass transfer and initiates a CE phase combined with the optimistic treatment for CE. This model leads to the highest predictions for the detection's among all variations that we have considered.

\subsubsection{Supernovae and compact remnants}

The formation of compact remnants and and their associated natal kicks also constitute important uncertainties.

In model \modRapid{} we consider the so-called ``rapid'' remnant mass function by \citet{Fryer+2015} as an alternative to the ``delayed'' description used in our detailed simulations. This affects the mass distribution (see Section~\ref{XX}), but the effects on the number of detectable systems is modest for BHBH and BHNS, and negligible for NSNS.

The same is true for changing the impact of changing the assumed maximum neutron star mass, $m_{\rm ns}$ (\modNSLow{} and \modNSHigh{}). Lowering $m_{\rm ns}$  increases the number of detections involving BHs and vice versa, but has no significant effect on the number of NSNS detections.

In Model \modNoPISN{}, we ignore pair instability supernovae, which has no significant effect on the predictions for the number of detections with LISA. This is not a surprise as the typical metallicity of the progenitors in our Milky Way is rather high.  The detectable systems are dominated by systems with final core masses that are well below the mass regime for pair instability supernovae.

Decreasing the natal kicks for all core-collapse supernovae (models \modSigLow{}-\modSigLower{}) increases the detection rates for each DCO type since lower kicks result in fewer disrupted binaries and hence a more numerous detectable population. The NSNS systems are most strongly affected.   The impact on BHNS systems is mild and even insignificant for BHBH systems. This is because, in our models, the natal kicks for BHs are scaled down with the amount of mass that falls back, which is large in the case of BHs.

In model \modNoBH{} we assume BHs form without any kick, while using our fiducial assumption for the natal kicks of neutron stars.  This increases the predictions for BHNS by a factor 3. The impact on BHBH systems is much smaller, This is because they mostly form from more massive cores which are already assumed to form with low kicks in our simulations.  As expected, the NSNS population is not affected.

%We note that the these two models best match the detection rates from LIGO (Broekgaarden et al.\ subm). Therefore we find the best matching models for the LIGO detection rates also produce the highest LISA detection rates.

\subsubsection{Stellar winds}
Mass loss in the form of stellar winds or eruptions are also a main incertainty. It affects by how much stars can grow in size and it affects the final core masses.

\citet{Broekgaarden+2021} provides us with model variations that consider variations in the mass loss by hot and luminous Wolf-Rayet stars.   We find that a reduction of the wind mass loss  has very little effect on our predictions.  It turns out that this is the consequence of several effects that cancel each other.

Firstly, the decreased winds mean that the DCOs (particularly those containing BHs) are generally more massive and so more detectable in LISA. Secondly, one may expect that LISA sources in model \modWRLow{} would be higher frequency than in our fiducial model as decreased winds generally result in tighter binaries. However, though this is the case at DCO formation, we find by the time the sources have evolved until the LISA mission they are actually at lower frequencies than in our fiducial model. This is because the reduced winds allow DCOs to be formed at higher metallicity and, therefore, at more recent times. This means that the DCOs do not evolve for as long before the LISA mission and so remain at lower frequencies (wider separations) thus making them less detectable.

In addition, we find that NSNSs are more eccentric and BHBHs are less eccentric than our fiducial model (with BHNSs relatively unchanged). The increase in eccentricity for NSNSs comes from the same reason as the lower frequency, more recent birth times mean that binaries have less time to circularise. The same is not true for BHBHs as the more massive systems are less affected by supernova kicks and so fewer high eccentricity systems are formed. Overall, despite the large differences in the system properties, these three effects in combination leave the detection rates relatively unchanged.

In model \modWRHigh{} we instead \textit{increase} the efficiency of Wolf-Rayet winds by a factor of 5. We note that there is currently no observational support for such a strong increase in the stellar winds of Wolf-Rayet stars, but it is nevertheless still interesting to consider.  In this model the detection rate of BHBHs decreases by over a factor of 10 and BHNSs by over a factor of 2 whilst the NSNS rate is relatively unchanged. Increasing the efficiency of WR winds strongly decreases the final masses of DCOs. This means that many progenitors that would have formed LISA sources under our fiducial assumptions would not have enough mass to produce a DCO, or produce a NS instead of a BH (therefore this effect is stronger on BHBHs than NSNSs). Moreover, DCOs that are formed tend to be less massive and therefore less detectable.

\subsection{Properties of detectable systems}\label{sec:property_variations}

In this section, we consider how varying underlying physics assumptions changes the properties of detectable systems. We focus on several key differences across physics variations rather than showing the differences in every model and thus this section is by no means exhaustive.

\subsubsection{Using LISA to investigate the lower mass gap}\label{sec:lower_mass_gap}

\begin{figure}[tb]
    \centering
    \includegraphics[width=\columnwidth]{fig10_lower_mass_gap_rapid_variation.pdf}
    \caption{Comparison of the component mass distribution of LISA detectable DCOs when using the Fryer delayed (model \modFid{}) and rapid (model \modRapid{}) remnant mass prescriptions. Distributions are plotted in the same way as Fig.~\ref{fig:fiducial_pdf_distributions}, except all DCOs types are shown in one curve and each type is weighted by its detection rate in the respective model. \href{https://github.com/TomWagg/detecting-DCOs-in-LISA/blob/main/paper/figures/fig10_lower_mass_gap_rapid_variation.pdf}{\faFileImage} \href{https://github.com/TomWagg/detecting-DCOs-in-LISA/blob/main/paper/figure_notebooks/variations.ipynb}{\faBook}.}
    \label{fig:lower_mass_gap_variation}
\end{figure}

In Figure~\ref{fig:lower_mass_gap_variation}, we show the component mass distribution for all LISA detectable DCOs (BHBHs, BHNSs and NSNSs) for two different remnant mass prescriptions. The grey distribution uses the Fryer \textit{delayed} remnant mass prescription \citep{Fryer+2012}, which is our fiducial assumption (model \modFid{}). This prescription produces compact objects in the lower mass gap ($2.5 \unit{M_{\odot}} \le m \le 5 \unit{M_{\odot}}$) and indeed we find that, of the LISA detectable DCOs, approximately \BHBHatLeastOneLowerMassGapPerc{}\% of BHBHs, \BHNSatLeastOneLowerMassGapPerc{}\% of BHNSs and \NSNSatLeastOneLowerMassGapPerc{}\% of NSNSs have at least one component in the lower mass gap. Overall, weighting by the relative detection rates, this gives that 55\% of our predicted LISA DCO detections would have at least one component in the lower mass gap when using this remnant mass prescription.

Alternatively, the blue curve in Fig.~\ref{fig:lower_mass_gap_variation} shows the same distribution but for the \textit{rapid} remnant mass prescription \citep{Fryer+2012}, which we use in model \modRapid{}. In this case, no compact objects are formed (and therefore, detected) in the lower mass gap. From the stark difference between these models, it is clear that it is difficult at this point to say with any certainty what fraction of systems LISA will detect in the lower mass gap given the highly uncertain formation rate of systems in this mass range.

However, it is important to highlight that \textit{if} DCOs are formed with components in the lower mass gap, LISA \textit{will} be able to detect them. And thus, LISA could be a useful instrument for providing evidence for the existence or non-existence of a lower mass gap based on the mass distribution of detected DCOs.

\subsubsection{Effect of natal kicks on eccentricity distribution}

\begin{figure}[tb]
    \centering
    \includegraphics[width=\columnwidth]{fig11_ecc_low_kick_variation.pdf}
    \caption{As Fig.~\ref{fig:fiducial_pdf_distributions}d, but for model \modSigLower{}. For comparison, we show the mean distribution for the fiducial model (model \modFid{}) as dashed lines. \href{https://github.com/TomWagg/detecting-DCOs-in-LISA/blob/main/paper/figures/fig11_ecc_low_kick_variation.pdf}{\faFileImage} \href{https://github.com/TomWagg/detecting-DCOs-in-LISA/blob/main/paper/figure_notebooks/variations.ipynb}{\faBook}.}
    \label{fig:ecc_low_kick_variation}
\end{figure}

In Figure~\ref{fig:ecc_low_kick_variation}, we investigate how decreasing the magnitude of natal kicks from core-collapse supernovae affects the eccentricity distribution of LISA detectable DCOs. For reference, we show the mean fiducial distributions (model \modFid{}) as dashed lines (see Fig.~\ref{fig:fiducial_pdf_distributions}d for full comparison). In the main curves, we reduce the velocity dispersion for core-collapse supernovae from $265 \unit{km}{s^{-1}}$ to $30 \unit{km}{s^{-1}}$ (model \modSigLower{}).

We find that the LISA detectable BHBHs are significantly less eccentric with weaker kicks, such that the population above $e = 0.2$ is nearly completely eliminated. This is because BHBHs are often massive enough to withstand strong natal kicks without disrupting and these kicks tend to impart significant eccentricity. In model \modSigLower{}, very few systems are ever given such strong kicks and thus very few BHBHs are detected with significant eccentricity.

Since BHNSs are less massive than BHBHs and have more unequal mass ratios, they are more vulnerable to disruption during supernova kicks. BHNSs can only withstand strong kicks when they are aimed in the correct direction and so only a small `lucky' fraction of the fiducial population is highly eccentric. Therefore in model \modSigLower{}, although we see that the population of highly eccentric BHNS systems is eliminated (similar to BHBHs), the peak of the distribution actually shifts to \textit{higher} eccentricity. This is because a larger fraction of systems are given weaker kicks that BHNSs can withstand and these impart much more moderate eccentricities.

Finally, we find that the NSNS distribution is relatively unchanged between model \modFid{} and \modSigLower{}. This is not surprising however since the majority of NSNSs are formed through electron-capture supernovae and ultra-stripped supernovae and for these types of supernovae we use $\sigma^{\rm 1D}_{\rm rms} = 30 \unit{km}{s^{-1}}$ already (see App.~\ref{app:fiducial_physics}) and thus there is little difference between the models.

Overall, we find that decreasing supernova natal kicks, though it strongly increases the number of detections (see Fig.~\ref{fig:detection_rates}), strongly \textit{decreases} the fraction of highly eccentric systems that are detected.

\subsubsection{Effect of Wolf-Rayet winds on mass distribution}

\begin{figure}[tb]
    \centering
    \includegraphics[width=\columnwidth]{fig12_wr_wind_mass_variations.pdf}
    \caption{As Fig.~\ref{fig:lower_mass_gap_variation}, but instead varying Wolf-Rayet wind efficiency (models \modWRLow{} and \modWRHigh{}). The curve for $f_{\rm WR} = 5$ has much higher uncertainties as there are many fewer systems formed in this model (the inverse reasoning also explains the lower uncertainties for $f_{\rm WR} = 0.1$). \href{https://github.com/TomWagg/detecting-DCOs-in-LISA/blob/main/paper/figures/fig12_wr_wind_mass_variations.pdf}{\faFileImage} \href{https://github.com/TomWagg/detecting-DCOs-in-LISA/blob/main/paper/figure_notebooks/variations.ipynb}{\faBook}.}
    \label{fig:wr_wind_mass_variations}
\end{figure}

In Fig.~\ref{fig:wr_wind_mass_variations} we show the effect that changing the efficiency of Wolf-Rayet winds has on the individual component mass distribution. Decreasing the Wolf-Rayet wind efficiency allows the formation of more massive DCOs in the Milky Way and, indeed, we see that the distribution extends to $25 \unit{M_{\odot}}$ and fewer detectable systems are formed at low masses. By contrast, increasing the Wolf-Rayet efficiency by a factor of 5 strongly disfavours the formation of systems at high masses and approximately 85\% of detectable systems have masses below $5 \unit{M_{\odot}}$. These three distributions are very distinct and so it is possible that the mass distribution of LISA could help to constrain the efficiency of Wolf-Rayet winds.