### 2021-11-16 paper/tex/paper.tex
most comments up to variations resolved

\documentclass[twocolumn, twocolappendix]{aastex63}
\received{\today}
\shorttitle{Detecting DCOs with LISA}
\graphicspath{{../figures/}}

\usepackage{lipsum}
\usepackage{physics}
\usepackage{multirow}
\usepackage{xspace}
\usepackage{natbib}
\usepackage{fontawesome5}
\usepackage{xcolor}
\usepackage{wrapfig}
\usepackage[figuresright]{rotating}

% remove indents in footnotes
\usepackage[hang,flushmargin]{footmisc}

\newcommand{\todo}[1]{{\color{red}{[TODO: #1}]}}
\newcommand{\tom}[1]{\textcolor{ForestGreen}{[Tom: #1]}}
\newcommand{\floor}[1]{\textcolor{Bittersweet}{[Floor: #1]}}
% custom function for adding units
\makeatletter
\newcommand{\unit}[1]{%
    \,\mathrm{#1}\checknextarg}
\newcommand{\checknextarg}{\@ifnextchar\bgroup{\gobblenextarg}{}}
\newcommand{\gobblenextarg}[1]{\,\mathrm{#1}\@ifnextchar\bgroup{\gobblenextarg}{}}
\makeatother

\newcommand{\Rsun}{\ensuremath{\,\rm{R_\odot}}\xspace}
\newcommand{\Msun}{\ensuremath{\,\rm{M}_{\odot}}\xspace}

\newcommand{\avg}[1]{\left\langle#1\right\rangle}

%% Physics variations shortcuts
\newcommand{\modFid}{A}
\newcommand{\modBetaLow}{B}
\newcommand{\modBetaMed}{C}
\newcommand{\modBetaHigh}{D}
\newcommand{\modCaseBB}{E}
\newcommand{\modCaseBBOpt}{F}
\newcommand{\modAlphaLowest}{G}
\newcommand{\modAlphaLow}{H}
\newcommand{\modAlphaHigh}{I}
\newcommand{\modAlphaHighest}{J}
\newcommand{\modOpt}{K}
\newcommand{\modRapid}{L}
\newcommand{\modNSLow}{M}
\newcommand{\modNSHigh}{N}
\newcommand{\modNoPISN}{O}
\newcommand{\modSigLow}{P}
\newcommand{\modSigLower}{Q}
\newcommand{\modNoBH}{R}
\newcommand{\modWRLow}{S}
\newcommand{\modWRHigh}{T}

\newcommand{\modRangeMT}{B-D}
\newcommand{\modRangeCE}{E-K}
\newcommand{\modRangeSN}{L-R}
\newcommand{\modRangeML}{S-T}

\newcommand{\nModels}{20}
\newcommand{\nMinusOneModels}{19}

\newcommand{\BHBHFourYear}{74}
\newcommand{\BHNSFourYear}{42}
\newcommand{\NSNSFourYear}{8}
\newcommand{\BHBHTenYear}{118}
\newcommand{\BHNSTenYear}{71}
\newcommand{\NSNSTenYear}{13}

% percentage of systems in lower mass gap
\newcommand{\BHBHLowerMassGap}{35\%}
\newcommand{\BHNSLowerMassGap}{39\%}

% percentage of systems with q > 0.8
\newcommand{\BHBHqAbovePointEight}{8\%}
\newcommand{\BHNSqAbovePointEight}{0\%}
\newcommand{\NSNSqAbovePointEight}{90\%}

% percentage of systems with e > 0.01
\newcommand{\BHBHNotCirc}{87\%}
\newcommand{\BHNSNotCirc}{46\%}
\newcommand{\NSNSNotCirc}{86\%}

% percentage of systems with e > 0.3
\newcommand{\BHBHHighlyEccentric}{21\%}
\newcommand{\BHNSHighlyEccentric}{8\%}
\newcommand{\NSNSHighlyEccentric}{12\%}

% number of systems with fractional chirp mass uncertainty < 1
\newcommand{\BHBHChirpMassMeasureableFour}{20}
\newcommand{\BHNSChirpMassMeasureableFour}{17}
\newcommand{\NSNSChirpMassMeasureableFour}{4}
\newcommand{\BHBHChirpMassMeasureableTen}{52}
\newcommand{\BHNSChirpMassMeasureableTen}{44}
\newcommand{\NSNSChirpMassMeasureableTen}{9}

% percentage of systems above Chandrasekhar mass
\newcommand{\BHBHAboveMaxWDWDFourPerc}{24}
\newcommand{\BHNSAboveMaxWDWDFourPerc}{28}
\newcommand{\NSNSAboveMaxWDWDFourPerc}{4}
\newcommand{\BHBHAboveMaxWDWDTenPerc}{38}
\newcommand{\BHNSAboveMaxWDWDTenPerc}{41}
\newcommand{\NSNSAboveMaxWDWDTenPerc}{5}

% percentage of systems with something other than 1 detectable harmonic
\newcommand{\BHBHMultipleHarmonicsFourPerc}{55}
\newcommand{\BHNSMultipleHarmonicsFourPerc}{27}
\newcommand{\NSNSMultipleHarmonicsFourPerc}{66}
\newcommand{\BHBHMultipleHarmonicsTenPerc}{61}
\newcommand{\BHNSMultipleHarmonicsTenPerc}{29}
\newcommand{\NSNSMultipleHarmonicsTenPerc}{68}

% percentage of systems with >1 harmonic and in disc
\newcommand{\BHBHEccInDiscFourPerc}{40}
\newcommand{\BHNSEccInDiscFourPerc}{23}
\newcommand{\NSNSEccInDiscFourPerc}{59}
\newcommand{\BHBHEccInDiscTenPerc}{40}
\newcommand{\BHNSEccInDiscTenPerc}{23}
\newcommand{\NSNSEccInDiscTenPerc}{59}

% number of systems that are definitely not WDWDs
\newcommand{\BHBHNotWDWDFour}{37}
\newcommand{\BHNSNotWDWDFour}{18}
\newcommand{\NSNSNotWDWDFour}{5}
\newcommand{\BHBHNotWDWDTen}{70}
\newcommand{\BHNSNotWDWDTen}{38}
\newcommand{\NSNSNotWDWDTen}{8}

% number of systems that can be determined as either a BH or NS
\newcommand{\BHBHEitherBHOrNSFour}{16}
\newcommand{\BHNSEitherBHOrNSFour}{7}
\newcommand{\NSNSEitherBHOrNSFour}{4}
\newcommand{\BHBHEitherBHOrNSTen}{39}
\newcommand{\BHNSEitherBHOrNSTen}{17}
\newcommand{\NSNSEitherBHOrNSTen}{8}
\newcommand{\BHBHEitherBHOrNSFourPerc}{21}
\newcommand{\BHNSEitherBHOrNSFourPerc}{18}
\newcommand{\NSNSEitherBHOrNSFourPerc}{47}
\newcommand{\BHBHEitherBHOrNSTenPerc}{33}
\newcommand{\BHNSEitherBHOrNSTenPerc}{24}
\newcommand{\NSNSEitherBHOrNSTenPerc}{62}

\newcommand{\BHBHatLeastOneLowerMassGapPerc}{69}
\newcommand{\BHNSatLeastOneLowerMassGapPerc}{39}
\newcommand{\NSNSatLeastOneLowerMassGapPerc}{0}
\newcommand{\ALLatLeastOneLowerMassGapPerc}{55}

\newcommand{\confinv}[3]{$#1${\raisebox{0.5ex}{\tiny$_{-#2}^{+#3}$}}}
\newcommand{\boldconfinv}[3]{$\mathbf{#1}${\raisebox{0.5ex}{\tiny$\mathbf{_{-#2}^{+#3}}$}}}

\definecolor{Blush}{rgb}{0.87, 0.36, 0.51}
\newcommand{\SdM}[1]{{\color{Blush}{#1}}}

% Milky Way notation
\newcommand{\achem}{\ensuremath{[\alpha/\mathrm{Fe}]}}

\begin{document}

\title{{\large Gravitational wave sources in our Galactic backyard:}\\{\normalsize Models for detecting BHBH, BHNS and NSNS binaries with LISA}}

% affiliations
\newcommand{\cfa}{Center for Astrophysics | Harvard \& Smithsonian, 60 Garden Street, Cambridge, MA 02138, USA}
\newcommand{\mpa}{Max-Planck-Institut für Astrophysik, Karl-Schwarzschild-Straße 1, 85741 Garching, Germany}
\newcommand{\cca}{Center for Computational Astrophysics, Flatiron Institute, 162 Fifth Ave, New York, NY, 10010, USA}
\newcommand{\UvA}{Anton Pannekoek Institute for Astronomy \& Grappa, University of Amsterdam, Postbus 94249, 1090 GE Amsterdam, The Netherlands}
\newcommand{\UW}{Department of Astronomy, University of Washington, Seattle, WA, 98195}

\author[0000-0001-6147-5761]{T. Wagg}
\affiliation{\UW}
\affiliation{\cfa}
\affiliation{\mpa}

\author[0000-0002-4421-4962]{F.S. Broekgaarden}
\affiliation{\cfa}

\author[0000-0001-9336-2825]{S.E. de Mink}
\affiliation{\mpa}
\affiliation{\UvA}
\affiliation{\cfa}

\author[0000-0002-6411-8695]{N. Frankel}
\affiliation{Max Planck Institute for Astronomy, Königstuhl 17, D-69117 Heidelberg, Germany}

\author[0000-0001-5484-4987]{L.A.C. van Son}
\affiliation{\cfa}
\affiliation{\UvA}
\affiliation{\mpa}

\author[0000-0001-7969-1569]{S. Justham}
\affiliation{School of Astronomy \& Space Science, University of the Chinese Academy of Sciences, Beijing 100012, China}
\affiliation{\UvA}
\affiliation{\mpa}

\correspondingauthor{Tom Wagg}
\email{tomjwagg@gmail.com}

\begin{abstract}
    \input{sections/abstract.tex}
\end{abstract}

\keywords{gravitational waves, LISA, black hole, neutron star, binary stars}

 \section{Introduction} \label{sec:intro}
\input{sections/introduction.tex}

\section{Method} \label{sec:method}
\input{sections/methods.tex}

\section{Results I - Predictions for LISA sources} \label{sec:results}
\input{sections/results.tex}

\section{Results II - Impact of physics assumptions} \label{sec:variations}
\input{sections/variations.tex}

\section{Discussion} \label{sec:discussion}
\input{sections/discussion.tex}

\section{Comparison with previous studies}\label{sec:compare_studies}
\input{sections/comparisons.tex}

\section{Conclusion \& Summary} \label{sec:conclusion}
\input{sections/conclusion.tex}

\acknowledgments{We thank Katie Brevik, Will Farr, Rob Farmer, Valeriya Korol, Floris Kummer, Eva Laplace, Mike Lau, Tyson Littenberg, Ilya Mandel, Mathieu Renzo, Alberto Sesana, Katie Sharpe, Simon Stevenson, the CCA GW group and COMPAS collaboration for insightful discussions. TW also thanks Terence Lovatt for his advice and help on improving an earlier draft of this project. Portions of this study were inspired by the LISA Sprint at the Center for Computational Astrophysics of the Flatiron Institute, supported by the Simons Foundation. This project was funded in part by the National Science Foundation under Grant No. (NSF grant number 2009131), the European Union’s Horizon 2020 research and innovation program from the European Research Council (ERC, Grant agreement No. 715063), and by the Netherlands Organization for Scientific Research (NWO) as part of the Vidi research program BinWaves with project number 639.042.728. We further acknowledge the Black Hole Initiative funded by a generous contribution of the John Templeton Foundation and the Gordon and Betty Moore Foundation.}

\software{We used LEGWORK to evolve sources over time and calculate signal-to-noise ratios, it is freely available at \url{https://legwork.readthedocs.io/en/latest/}. Simulations in this paper made use of the COMPAS rapid binary population synthesis code, which is freely available at \url{http://github.com/TeamCOMPAS/COMPAS} \citep{Stevenson+2017, Vigna-Gomez+2018, Broekgaarden+2019}. The simulations performed in this work were simulated with a COMPAS version that predates the publicly available code. Our version of the code is most similar to version 02.13.01 of the publicly available COMPAS code. Requests for the original code can be made to Floor Broekgaarden. The authors used {\sc{STROOPWAFEL}} from \citep{Broekgaarden+2019}, publicly available at \url{https://github.com/FloorBroekgaarden/STROOPWAFEL}\footnote{For the latest pip installable version of STROOPWAFEL please contact Floor Broekgaarden.}.
The authors also made use of Python from the Python Software Foundation. Python Language Reference, version 3.6. Available at \url{http://www.python.org}. In addition the following Python packages were used: matplotlib (\url{https://matplotlib.org/}), {NumPy} (\url{https://numpy.org/}), SciPy (\url{https://www.scipy.org/}), Jupyter Lab (\url{https://jupyter.org/}), seaborn (\url{https://seaborn.pydata.org/}), h5py (https://www.h5py.org/) and Astropy (\url{http://www.astropy.org}).  This research has made use of NASA’s Astrophysics Data System Bibliographic Services. We also made use of the computational facilities from the Harvard FAS Research Computing cluster.}

\clearpage

\bibliographystyle{aasjournal}
\bibliography{references}

\restartappendixnumbering

\allowdisplaybreaks
\appendix
\input{sections/appendix.tex}

\end{document}