### 2021-11-26 paper/tex/paper.tex
Updates from Overleaf

\documentclass[twocolumn, twocolappendix]{aastex63}
\received{\today}
\shorttitle{Detecting DCOs with LISA}
\graphicspath{{../figures/}}

\usepackage{lipsum}
\usepackage{physics}
\usepackage{multirow}
\usepackage{xspace}
\usepackage{natbib}
\usepackage{fontawesome5}
\usepackage{xcolor}
\usepackage{wrapfig}
\usepackage[figuresright]{rotating}

% remove indents in footnotes
\usepackage[hang,flushmargin]{footmisc}

\newcommand{\todo}[1]{{\color{red}{[TODO: #1}]}}
\newcommand{\tom}[1]{\textcolor{ForestGreen}{[Tom: #1]}}
\newcommand{\floor}[1]{\textcolor{Bittersweet}{[Floor: #1]}}
% custom function for adding units
\makeatletter
\newcommand{\unit}[1]{%
    \,\mathrm{#1}\checknextarg}
\newcommand{\checknextarg}{\@ifnextchar\bgroup{\gobblenextarg}{}}
\newcommand{\gobblenextarg}[1]{\,\mathrm{#1}\@ifnextchar\bgroup{\gobblenextarg}{}}
\makeatother

\newcommand{\Rsun}{\ensuremath{\,\rm{R_\odot}}\xspace}
\newcommand{\Msun}{\ensuremath{\,\rm{M}_{\odot}}\xspace}

\newcommand{\avg}[1]{\left\langle#1\right\rangle}

%% Physics variations shortcuts
\newcommand{\modFid}{A}
\newcommand{\modBetaLow}{B}
\newcommand{\modBetaMed}{C}
\newcommand{\modBetaHigh}{D}
\newcommand{\modCaseBB}{E$^{\prime}$}
\newcommand{\modCaseBBOpt}{F}
\newcommand{\modAlphaLowest}{G}
\newcommand{\modAlphaLow}{H}
\newcommand{\modAlphaHigh}{I}
\newcommand{\modAlphaHighest}{J}
\newcommand{\modOpt}{K}
\newcommand{\modRapid}{L}
\newcommand{\modNSLow}{M}
\newcommand{\modNSHigh}{N}
\newcommand{\modNoPISN}{O}
\newcommand{\modSigLow}{P}
\newcommand{\modSigLower}{Q}
\newcommand{\modNoBH}{R}
\newcommand{\modWRLow}{S}
\newcommand{\modWRHigh}{T}

\newcommand{\modRangeMT}{B-D}
\newcommand{\modRangeCE}{E$^{\prime}$-K}
\newcommand{\modRangeSN}{L-R}
\newcommand{\modRangeML}{S-T}

\newcommand{\nModels}{20}
\newcommand{\nMinusOneModels}{19}

\newcommand{\rangeFourYear}{30-370}
\newcommand{\rangeTenYear}{50-550}

\newcommand{\BHBHFourYear}{74}
\newcommand{\BHNSFourYear}{42}
\newcommand{\NSNSFourYear}{8}
\newcommand{\BHBHTenYear}{118}
\newcommand{\BHNSTenYear}{71}
\newcommand{\NSNSTenYear}{13}

% percentage of systems in lower mass gap
\newcommand{\BHBHLowerMassGap}{35\%}
\newcommand{\BHNSLowerMassGap}{39\%}

% percentage of systems with q > 0.8
\newcommand{\BHBHqAbovePointEight}{8\%}
\newcommand{\BHNSqAbovePointEight}{0\%}
\newcommand{\NSNSqAbovePointEight}{90\%}

% percentage of systems with e > 0.01
\newcommand{\BHBHNotCirc}{87\%}
\newcommand{\BHNSNotCirc}{46\%}
\newcommand{\NSNSNotCirc}{86\%}

% percentage of systems with e > 0.3
\newcommand{\BHBHHighlyEccentric}{21\%}
\newcommand{\BHNSHighlyEccentric}{8\%}
\newcommand{\NSNSHighlyEccentric}{12\%}

% number of systems with fractional chirp mass uncertainty < 1
\newcommand{\BHBHChirpMassMeasureableFour}{20}
\newcommand{\BHNSChirpMassMeasureableFour}{17}
\newcommand{\NSNSChirpMassMeasureableFour}{4}
\newcommand{\BHBHChirpMassMeasureableTen}{52}
\newcommand{\BHNSChirpMassMeasureableTen}{44}
\newcommand{\NSNSChirpMassMeasureableTen}{9}

% percentage of systems above Chandrasekhar mass
\newcommand{\BHBHAboveMaxWDWDFourPerc}{24}
\newcommand{\BHNSAboveMaxWDWDFourPerc}{28}
\newcommand{\NSNSAboveMaxWDWDFourPerc}{4}
\newcommand{\BHBHAboveMaxWDWDTenPerc}{38}
\newcommand{\BHNSAboveMaxWDWDTenPerc}{41}
\newcommand{\NSNSAboveMaxWDWDTenPerc}{5}

% percentage of systems with something other than 1 detectable harmonic
\newcommand{\BHBHMultipleHarmonicsFourPerc}{55}
\newcommand{\BHNSMultipleHarmonicsFourPerc}{27}
\newcommand{\NSNSMultipleHarmonicsFourPerc}{66}
\newcommand{\BHBHMultipleHarmonicsTenPerc}{61}
\newcommand{\BHNSMultipleHarmonicsTenPerc}{29}
\newcommand{\NSNSMultipleHarmonicsTenPerc}{68}

% percentage of systems with >1 harmonic and in disc
\newcommand{\BHBHEccInDiscFourPerc}{40}
\newcommand{\BHNSEccInDiscFourPerc}{23}
\newcommand{\NSNSEccInDiscFourPerc}{59}
\newcommand{\BHBHEccInDiscTenPerc}{40}
\newcommand{\BHNSEccInDiscTenPerc}{23}
\newcommand{\NSNSEccInDiscTenPerc}{59}

% number of systems that are definitely not WDWDs
\newcommand{\BHBHNotWDWDFour}{37}
\newcommand{\BHNSNotWDWDFour}{18}
\newcommand{\NSNSNotWDWDFour}{5}
\newcommand{\BHBHNotWDWDTen}{70}
\newcommand{\BHNSNotWDWDTen}{38}
\newcommand{\NSNSNotWDWDTen}{8}

% number of systems that can be determined as either a BH or NS
\newcommand{\BHBHEitherBHOrNSFour}{16}
\newcommand{\BHNSEitherBHOrNSFour}{7}
\newcommand{\NSNSEitherBHOrNSFour}{4}
\newcommand{\BHBHEitherBHOrNSTen}{39}
\newcommand{\BHNSEitherBHOrNSTen}{17}
\newcommand{\NSNSEitherBHOrNSTen}{8}
\newcommand{\BHBHEitherBHOrNSFourPerc}{21}
\newcommand{\BHNSEitherBHOrNSFourPerc}{18}
\newcommand{\NSNSEitherBHOrNSFourPerc}{47}
\newcommand{\BHBHEitherBHOrNSTenPerc}{33}
\newcommand{\BHNSEitherBHOrNSTenPerc}{24}
\newcommand{\NSNSEitherBHOrNSTenPerc}{62}

\newcommand{\BHBHatLeastOneLowerMassGapPerc}{69}
\newcommand{\BHNSatLeastOneLowerMassGapPerc}{39}
\newcommand{\NSNSatLeastOneLowerMassGapPerc}{0}
\newcommand{\ALLatLeastOneLowerMassGapPerc}{55}

\newcommand{\citefloorp}{(\citealt{Broekgaarden+2021}, Broekgaarden et al.\ (in prep.))}

\newcommand{\confinv}[3]{$#1${\raisebox{0.5ex}{\tiny$_{-#2}^{+#3}$}}}
\newcommand{\boldconfinv}[3]{$\mathbf{#1}${\raisebox{0.5ex}{\tiny$\mathbf{_{-#2}^{+#3}}$}}}

\definecolor{Blush}{rgb}{0.87, 0.36, 0.51}
\newcommand{\SdM}[1]{{\color{Blush}{#1}}}

% Milky Way notation
\newcommand{\achem}{\ensuremath{[\alpha/\mathrm{Fe}]}}

\begin{document}

\title{{Gravitational wave sources in our Galactic backyard:}\\{Predictions for BHBH, BHNS and NSNS binaries detectable with LISA}}

% affiliations
\newcommand{\cfa}{Center for Astrophysics | Harvard \& Smithsonian, 60 Garden Street, Cambridge, MA 02138, USA}
\newcommand{\mpa}{Max-Planck-Institut für Astrophysik, Karl-Schwarzschild-Straße 1, 85741 Garching, Germany}
\newcommand{\cca}{Center for Computational Astrophysics, Flatiron Institute, 162 Fifth Ave, New York, NY, 10010, USA}
\newcommand{\UvA}{Anton Pannekoek Institute for Astronomy \& Grappa, University of Amsterdam, Postbus 94249, 1090 GE Amsterdam, The Netherlands}
\newcommand{\UW}{Department of Astronomy, University of Washington, Seattle, WA, 98195}

\author[0000-0001-6147-5761]{T. Wagg}
\affiliation{\UW}
\affiliation{\cfa}
\affiliation{\mpa}

\author[0000-0002-4421-4962]{F.S. Broekgaarden}
\affiliation{\cfa}

\author[0000-0001-9336-2825]{S.E. de Mink}
\affiliation{\mpa}
\affiliation{\UvA}
\affiliation{\cfa}

\author[0000-0002-6411-8695]{N. Frankel}
\affiliation{Canadian Institute for Theoretical Astrophysics, University of Toronto, 60 St. George Street, Toronto, ON M5S 3H8, Canada}
\affiliation{Max Planck Institute for Astronomy, K\"onigstuhl 17, D-69117 Heidelberg, Germany}
\affiliation{David A. Dunlap Department for Astronomy and Astrophysics, University
of Toronto, 50 St. George Street, Toronto, ON M5S 3H4, Canada}

\author[0000-0001-5484-4987]{L.A.C. van Son}
\affiliation{\cfa}
\affiliation{\UvA}
\affiliation{\mpa}

\author[0000-0001-7969-1569]{S. Justham}
\affiliation{School of Astronomy \& Space Science, University of the Chinese Academy of Sciences, Beijing 100012, China}
\affiliation{\UvA}
\affiliation{\mpa}

\correspondingauthor{Tom Wagg}
\email{tomjwagg@gmail.com}

\begin{abstract}
    \input{sections/abstract.tex}
\end{abstract}

\keywords{gravitational waves, gravitational wave detectors, compact objects, stellar mass black holes, neutron stars, binary stars, stellar evolution, pulsars}

\section{Introduction} \label{sec:intro}
\input{sections/introduction.tex}

\section{Method} \label{sec:method}
\input{sections/methods.tex}

\section{Results I - Predictions for LISA sources} \label{sec:results}
\input{sections/results.tex}

\section{Results II - Impact of physics assumptions} \label{sec:variations}
\input{sections/variations.tex}

\section{Discussion} \label{sec:discussion}
\input{sections/discussion.tex}

\section{Comparison with previous studies}\label{sec:compare_studies}
\input{sections/comparisons.tex}

\section{Conclusion \& Summary} \label{sec:conclusion}
\input{sections/conclusion.tex}

\acknowledgments{We thank Katie Brevik, Will Farr, Rob Farmer, Valeriya Korol, Floris Kummer, Eva Laplace, Mike Lau, Tyson Littenberg, Ilya Mandel, Mathieu Renzo, Alberto Sesana, Katie Sharpe, Simon Stevenson, the CCA GW group and COMPAS collaboration for insightful discussions. TW also thanks Terence Lovatt for his advice and help on improving an earlier draft of this project. Portions of this study were inspired by the LISA Sprint at the Center for Computational Astrophysics of the Flatiron Institute, supported by the Simons Foundation. This project was funded in part by the National Science Foundation under Grant No. (NSF grant number 2009131), the European Union’s Horizon 2020 research and innovation program from the European Research Council (ERC, Grant agreement No. 715063), and by the Netherlands Organization for Scientific Research (NWO) as part of the Vidi research program BinWaves with project number 639.042.728. We further acknowledge the Black Hole Initiative funded by a generous contribution of the John Templeton Foundation and the Gordon and Betty Moore Foundation.}

\software{We used LEGWORK to evolve sources over time and calculate signal-to-noise ratios \citep{Wagg+2021}. It is freely available at \url{https://legwork.readthedocs.io/en/latest/}. Simulations in this paper made use of the COMPAS rapid binary population synthesis code \citep{COMPAS:2021methodsPaper}. The simulations performed in this work were simulated with a COMPAS version that predates the publicly available code. Our version is most similar to v02.13.01 of the publicly available COMPAS code. Requests for the original code can be made to Floor Broekgaarden. The authors used {\sc{STROOPWAFEL}} from \citep{Broekgaarden+2019}, publicly available at \url{https://github.com/FloorBroekgaarden/STROOPWAFEL}\footnote{For the latest pip installable version of STROOPWAFEL please contact Floor Broekgaarden.}.
The authors also made use of Python (v3.8), available at \url{http://www.python.org}. In addition the following Python packages were used: \texttt{matplotlib} \citep{Hunter+2007}, \texttt{NumPy} \citep{2020NumPy-Array}, \texttt{Astropy} \citep{AstropyCollaboration+2013,AstropyCollaboration+2018}, \texttt{Seaborn} \citep{Waskom+2021}, \texttt{SciPy} \citep{2020SciPy-NMeth}, \texttt{h5py} \citep{Collette+2021} and \texttt{Jupyter Lab} \citep{Kluyver+2016}. This research has made use of NASA’s Astrophysics Data System Bibliographic Services. We also made use of the computational facilities from the Harvard FAS Research Computing cluster.}

\bibliographystyle{aasjournal}
\bibliography{references}

\restartappendixnumbering

\allowdisplaybreaks
\appendix
\input{sections/appendix.tex}

\end{document}

% Oh no, I certainly hope no one sees this comment. Wouldn't it be awful if astro-ph-leaks tweeted this and advertised our paper (: